What does the crowd believe? A hierarchical approach to estimating subjective beliefs from empirical data

### 2016-05-11 text/01_CogSci_abstract/paper.tex
CogSci revisions and list of future steps

\usepackage{cogsci}
\usepackage{pslatex}
\usepackage{apacite}
\usepackage{subcaption}
\usepackage{todonotes}
\usepackage{amsmath}
\usepackage{color}
\usepackage{booktabs}
\usepackage{caption}
\captionsetup{belowskip=0pt,aboveskip=2pt}
\setlength{\textfloatsep}{9pt}

@@

\begin{abstract}
  People's beliefs about everyday events are both of theoretical interest in
  their own right and an important ingredient in model building---especially in 
  Bayesian cognitive models of phenomena such as logical reasoning, future predictions, and language
  use. Here, we explore several recently used methods for measuring subjective beliefs about
  unidimensional contiguous properties, such as the likely price of a new watch. We use a
  hierarchical Bayesian data-analysis model for inferring likely population-level beliefs as the central
  tendency of participants' individual-level beliefs.  Three different dependent measures are
  used to infer latent beliefs: (i) slider ratings of (relative) likelihood of intervals of
  values, (ii) a give-a-number task, and (iii) choice of the more likely of two intervals of
  values. Our results suggest that using averaged normalized slider ratings for binned
  quantities is a practical and fairly good approximator of inferred population-level beliefs.

@@

We cannot look into a person's head. But in making sense of observed behavior we readily
ascribe beliefs and desires to fellow agents. This happens intuitively, in folk psychology, but
also in science. Scientific ascription of latent mental states plays an important role in many
explanations of higher-order cognition: decision making, reasoning, language use, etc. It is
therefore vital to have methods for validating explanatory mental state ascriptions.

A family of models where this is particularly pressing are Bayesian models of cognition which
seek to explain task behavior in a variety of domains as partially informed by what
participants believe about mundane events---their \emph{prior beliefs}. Take interpretation of
language. 
``That watch cost a million dollars,'' tends to be interpreted as hyperbole: conveying affect rather than literal truth.
Empirical data on whether a statement like ``That watch cost $n$ dollars'' is
understood as hyperbole can be explained well by a
Bayesian model of utterance interpretation \citep{KaoWu2014:Nonliteral-Unde}. This model
assigns a crucial role to an empirical measure of participants' expectations about the likely
or normal price of a watch---only when the uttered price is sufficiently unlikely \emph{a priori} will hyperbolic interpretation be possible.  
Other examples of domains in which empirically successful models
have included a measure of participants' prior beliefs include making predictions about
everyday events \citep{GriffithsTenenbaum2006:Optimal-Predict}, referential communication
\citep{FrankGoodman2012:Predicting-Prag}, the strength of pragmatic enrichments
\citep{DegenTessler2015:Wonky-worlds:-L}, or the interpretation of quantifiers
\citep{SchollerFranke2015:Semantic-values}.

@@

More complex elicitation methods include \emph{scoring rules} \citep{Savage1971:Elicitation-of-,AndersenFountain2014:Estimating-Subj,SchlagTremewan2014:A-penny-for-you}, prominent in economics, and the \emph{iterated learning paradigm} \citep{LewandowskyGriffiths2009:The-Wisdom-of-I}.
These more complex methods are powerful but difficult to implement, and often assume very specific behavior from participants---such as optimal decision making under perfect knowledge of the payoff scheme.
%}. Scoring
%rules are complex schemes of assigning monetary rewards to participants' answers that make sure
%that responses (by rational participants) faithfully reflect a measure of interest: the actual
%subjective probability of an event, the mean of a distribution, a credible interval
%etc. Scoring-rule tasks, however, require participants to become sufficiently familiar with the (usually
%probabilistic) payoff scheme and assume close to optimal choices. 
%
%A final insightful, but also complex approach is to use a
%cleverly designed

@@

A simple technique that seems to fit this bill is the \emph{binned histogram} task that has
recently been used with apparent success
\cite<e.g.>{KaoWu2014:Nonliteral-Unde,KaoBergen2014:Formalizing--th,Tessler2015:Understanding-B,SchollerFranke2015:Semantic-values}. Participants
adjust sliders to express (relative) subjective beliefs about how likely it is that the value of an
uncertain contiguous quantity lies in some interval of possible values (a bin). E.g., to report
beliefs about the a watch's price, participants adjust sliders whose endpoints are labelled
``extremely likely'' and ``impossible.'' Each slider corresponds to one of 15 bins that
partition the range of plausible values (established by a pre-test), such as ``\$0-\$50'',
``\$50-\$100'', etc. up to ``\$700-\$750'' and ``more than \$750.'' Each participant's slider
ratings are normalized and the results averaged across participants. The resulting \emph{mean
  slider ratings} look like plausible population-level beliefs and give good results when fed
into cognitive models that predict task behavior based on prior expectations.

Explanatory success aside, the question remains whether mean slider ratings measure what we
would like them to. Are these good approximations of a central tendency of participants'
individual beliefs? Are these measures consistent with participants' behavior in other tasks,
such as \emph{give-a-number}? To address this issue we collected data in a
within-subject paradigm from three task types: (i) \emph{binned histograms}, (ii)
\emph{give-a-number} and (iii) \emph{paired comparisons}. The latter asks participants for a direct comparison of two bins
from the \emph{binned histograms} task, and was included as a further consistency check. 
We used a Bayesian hierarchical model to analyze this data jointly.
This model inferred likely
latent subjective beliefs, where each subjective belief is, intuitively, a noisy perturbation
of a population-level belief. We find that mean \emph{a posteriori} population-level beliefs
are approximated well by mean slider ratings from the binned-histograms task, suggesting that this may indeed be a
sound and easy measure of what the crowd believes.  
Moreover, the model is able to
capture participants' behavior in all three task types reasonably well, suggesting that what
mean slider ratings measure is what we think it is: a population-level central tendency of
latent subjective beliefs.

@@

\paragraph{Participants, procedure and materials} We recruited 20 self-reported English native
speakers over Mechanical Turk and collected responses for eight different items (listed in
\tableref{tab:Items}) using three different dependent measures: \emph{binned histograms},
\emph{give-a-number} and \emph{paired comparisons}. Each participant rated each item using each
dependent measure. Trial order was randomized.

@@

The data we would like to explain are: (i) the normalized slider ratings $s_{ijk} \in [0;1]$ of
participant $i \in \set{1, \dots, 20}$ for item $j \in \set{1, \dots, 8}$ and bin
$k \in \set{1, \dots, 15}$ from the binned histogram task; (ii) the bins $n_{ij} \in \set{1, \dots, 15}$ in which participant
$i$'s number choice for item $j$ was in the give-a-number task; and (iii) the binary choices $c_{ijl} \in \set{0,1}$ of
whether participant $i$ selected the higher bin for item $j$ in the paired comparison task for each comparison
$l$. There are two simplifications in need of commenting. In (i), we focus on slider ratings after
normalizing for each participant, because we assume that slider adjustments reflect relative, not
absolute estimates of subjective beliefs. In (ii), we focus on bin choices, not actual number
choices, in order to avoid, as much as possible, considerations of salience of particular
numbers, and also because otherwise data from items with smaller domains of plausible numbers
would get more weight than data from items which allow for a wider set of number choices.

@@

Of most theoretical interest are posterior estimates of $Q_j$. \figref{fig:PosteriorQj} shows
means of the posterior $Q_j$s, with their 95\% HDIs, alongside the mean slider ratings. The
latter provide a very good approximation of the latent, inferred population-level
beliefs. Inspection of posteriors of individual $P_{ij}$ shows that there is ample variation
between participants. Nonetheless, the way the model suggests we should think about harnessing
the individual $P_{ij}$s under a population-level central tendency is closely approximated by
mean slider ratings. Although the match is not perfect, it is good enough to say that the
latter are a practical way of approximating what the crowd believes despite individual
differences.

@@

Posterior predictive checks indicate that the trained model captures patterns of answers at the
aggregate population level well. To have a more fine-grained measure of model fit, we also
looked at posterior predictive $p$-values \citep{GelmanCarlin2014:Bayesian-Data-A} at the level
of participants and items. We focus on the binned histograms task, because here the
population-level posterior predictive checks revealed no systematic deviance. For this task,
fixing a participant and an item, observations and replicates are probability vectors of length
15. In a first analysis, we used the mean of these probability vectors as a test statistic. The
minimum posterior predictive $p$-value over all 20 (participants) times 8 (items) cases was
0.13, suggesting that the means of observed $s_{ij}$ are non-surprising to the trained
model. In a second analysis, we used entropy as a test statistic. Two cases gave posterior
predictive $p$-values lower than 0.05. These were from the two participants who gave a very
extreme slider rating for the ``marbles'' item, basically assigning all ``mass'' to the last
bin. What this suggests is that the model can cope reasonably well also with individual-level
data, but has problems accounting for ``extreme'' choices, given that the population-level
hyper-prior on $P_{ij}$ will lead to shrinkage.

@@

MF's work was supported by the Institutional Strategy of the University of T̈\"{u}bingen
(DFG, ZUK 63). MF, FD, AS and JD's work was supported by the
Priority Program XPrag.de (DFG Schwerpunktprogramm 1727). This work was further supported by
ONR grant N00014-13-1-0788 and a James S. McDonnell Foundation Scholar Award to NDG and an SNF
Early Postdoc.Mobility Award to JD.

>>>>

\usepackage{cogsci}
\usepackage{pslatex}
\usepackage{apacite}
\usepackage{subcaption}
\usepackage{todonotes}
\usepackage{amsmath}
\usepackage{color}
\usepackage{booktabs}
\usepackage{caption}
\captionsetup{belowskip=0pt,aboveskip=2pt}
\setlength{\textfloatsep}{8pt}

@@

\begin{abstract}
  People's beliefs about everyday events are both of theoretical interest in their own right
  and an important ingredient in model building---especially in Bayesian cognitive models of
  phenomena such as logical reasoning, future predictions, and language use. Here, we explore
  several recently used methods for measuring subjective beliefs about unidimensional
  contiguous properties, such as the likely price of a new watch. As a first step towards a way
  of assessing and comparing belief elicitation methods, we use hierarchical Bayesian modeling
  for inferring likely population-level beliefs as the central tendency of participants'
  individual-level beliefs. Three different dependent measures are considered: (i) slider
  ratings of (relative) likelihood of intervals of values, (ii) a give-a-number task, and (iii)
  choice of the more likely of two intervals of values. Our results suggest that using averaged
  normalized slider ratings for binned quantities is a practical and fairly good approximator
  of inferred population-level beliefs.

@@

When trying to understand observed behavior we readily ascribe beliefs and desires to fellow
agents. This happens intuitively, in folk psychology, but also in science. Scientific
ascription of latent mental states plays an important role in many explanations of higher-order
cognition: decision making, reasoning, language use, etc. It is therefore vital to have methods
for validating explanatory mental state ascriptions.

A family of models where this is particularly pressing are Bayesian models of cognition which
seek to explain task behavior in a variety of domains as partially informed by what
participants believe about mundane events---their \emph{prior beliefs}. Take interpretation of
language.  ``That watch cost a million dollars,'' tends to be interpreted as hyperbole:
conveying affect rather than literal truth.  Empirical data on whether similar statements are
understood as hyperbole can be explained well by a Bayesian model of utterance interpretation
\citep{KaoWu2014:Nonliteral-Unde}. This model assigns a crucial role to an empirical measure of
participants' expectations about the likely or normal price of a watch---only when the uttered
price is sufficiently unlikely \emph{a priori} will hyperbolic interpretation be possible.
Other examples of domains in which empirically successful models have included a measure of
participants' prior beliefs include making predictions about everyday events
\citep{GriffithsTenenbaum2006:Optimal-Predict}, referential reasoning
\citep{FrankGoodman2012:Predicting-Prag}, strength of pragmatic enrichments
\citep{DegenTessler2015:Wonky-worlds:-L}, or quantifier interpretation
\citep{SchollerFranke2015:Semantic-values}.

@@

More complex elicitation methods include \emph{scoring rules}
\citep{Savage1971:Elicitation-of-,AndersenFountain2014:Estimating-Subj,SchlagTremewan2014:A-penny-for-you},
prominent in economics, and the \emph{iterated learning paradigm}
\citep{LewandowskyGriffiths2009:The-Wisdom-of-I}.  These methods are powerful but difficult to
implement, and often assume very specific behavior from participants---such as optimal decision
making under perfect knowledge of the payoff scheme.
%}. Scoring
%rules are complex schemes of assigning monetary rewards to participants' answers that make sure
%that responses (by rational participants) faithfully reflect a measure of interest: the actual
%subjective probability of an event, the mean of a distribution, a credible interval
%etc. Scoring-rule tasks, however, require participants to become sufficiently familiar with the (usually
%probabilistic) payoff scheme and assume close to optimal choices. 
%
%A final insightful, but also complex approach is to use a
%cleverly designed

@@

A simple technique that seems to fit this bill is the \emph{binned histogram} task that has
recently been used with apparent success
\cite<e.g.>{KaoWu2014:Nonliteral-Unde,KaoBergen2014:Formalizing--th,Tessler2015:Understanding-B,SchollerFranke2015:Semantic-values}. Participants
adjust sliders to express (relative) subjective beliefs about how likely it is that the value
of an uncertain contiguous quantity lies in some interval of possible values (a bin). E.g., to
report beliefs about a watch's price, participants adjust sliders whose endpoints are labelled
``extremely likely'' and ``impossible.'' Each slider corresponds to one of 15 bins that
partition the range of plausible values (established by a pre-test), such as ``\$0-\$50'',
``\$50-\$100'', etc.~up to ``\$700-\$750'' and ``more than \$750.'' Each participant's slider
ratings are normalized and the results averaged across participants. The resulting \emph{mean
  slider ratings} look like plausible population-level beliefs and give good results when fed
into cognitive models that predict task behavior based on prior expectations.

Explanatory success aside, the question remains whether mean slider ratings measure what we
would like them to. Are these good approximations of a central tendency of participants'
individual beliefs? Are these measures consistent with participants' behavior in other tasks,
such as \emph{give-a-number}? To scrutinize the binned histogram task we address these issues
by collecting data in a within-subject paradigm from three task types: (i) \emph{binned
  histograms}, (ii) \emph{give-a-number} and (iii) \emph{paired comparisons}. The latter asks
participants for a direct comparison of two bins from the \emph{binned histograms} task, and
was included as a further consistency check.  We used a Bayesian hierarchical model to analyze
this data jointly.  The model infers latent subjective beliefs, where each subjective
belief is, intuitively, a noisy perturbation of a population-level belief. We find that mean
\emph{a posteriori} population-level beliefs are approximated well by mean slider ratings from
the binned-histograms task, suggesting that this may indeed be a sound and easy measure of what
the crowd believes.  Moreover, the model is able to capture participants' behavior in all three
task types reasonably well, suggesting that what mean slider ratings measure is what we think
it is: a population-level central tendency of latent subjective beliefs.

@@

\paragraph{Participants, procedure and materials} We recruited 20 self-reported English native
speakers over Mechanical Turk and collected responses for eight different items (listed in
\tableref{tab:Items}) using the three different dependent measures mentioned above. Each
participant rated each item using each dependent measure. Trial order was randomized.

@@

The data we would like to explain are: (i) the normalized slider ratings $s_{ijk} \in [0;1]$ of
participant $i \in \set{1, \dots, 20}$ for item $j \in \set{1, \dots, 8}$ and bin
$k \in \set{1, \dots, 15}$ from the binned histogram task; (ii) the bins
$n_{ij} \in \set{1, \dots, 15}$ in which participant $i$'s number choice for item $j$ was in
the give-a-number task; and (iii) the binary choices $c_{ijl} \in \set{0,1}$ of whether
participant $i$ selected the higher bin for item $j$ in the paired comparison task for each
comparison $l$. There are two simplifications in need of commenting. In (i), we focus on slider
ratings after normalizing for each participant, because we assume that slider adjustments
reflect relative, not absolute estimates of subjective beliefs. In (ii), we focus on bin
choices, not actual number choices, in order to avoid, as much as possible, considerations of
salience of particular numbers, and also because otherwise data from items with smaller domains
of plausible numbers would get more weight than data from items with a wider range of
number choices. (Future work should investigate the relation to an explicit \emph{give-a-bin}
task.)

@@

Posterior estimates of $Q_j$ are the most relevant. \figref{fig:PosteriorQj} shows their means
with their 95\% HDIs, alongside the mean slider ratings. The latter provide a very good
approximation of the inferred population-level beliefs. Inspection of posteriors of
individual $P_{ij}$ shows that there is ample variation between participants. Still, the
way the model suggests we should think about harnessing the individual $P_{ij}$s under a
population-level central tendency is closely approximated by mean slider ratings. Although the
match is not perfect, it is good enough to say that the latter are a practical way of
approximating what the crowd believes despite individual differences.

@@

Posterior predictive checks indicate that the trained model captures patterns of answers at the
aggregate population level well. To have a more fine-grained measure of model fit, we also
looked at posterior predictive $p$-values \citep{GelmanCarlin2014:Bayesian-Data-A} at the level
of participants and items. We look at the binned histograms task, because this is our main
focus here and population-level posterior predictive checks for BH revealed no systematic
deviance. Fixing a participant and an item, observations and replicates are probability vectors
of length 15. In a first analysis, we used the mean of these probability vectors as a test
statistic. The minimum posterior predictive $p$-value over all 20 (participants) times 8
(items) cases was 0.13, suggesting that the means of observed $s_{ij}$ are non-surprising to
the trained model. In a second analysis, we used entropy as a test statistic. Two cases gave
posterior predictive $p$-values lower than 0.05. These were from the two participants who gave
a very extreme slider rating for the ``marbles'' item, basically assigning all ``mass'' to the
last bin. What this suggests is that the model can cope reasonably well also with
individual-level data, but, somewhat unsurprisingly, has problems accounting for ``extreme''
choices, given that the population-level hyper-prior on $P_{ij}$ will lead to shrinkage.

@@

MF's work was supported by the Institutional Strategy of the University of T̈\"{u}bingen
(DFG, ZUK 63). MF, FD, AS and JD's work was supported by the
Priority Program XPrag.de (DFG Schwerpunktprogramm 1727). This work was further supported by
ONR grant N00014-13-1-0788 and a James S. McDonnell Foundation Scholar Award to NDG and an SNF
Early Postdoc.Mobility Award to JD. Thanks to three anonymous reviewers for helpful suggestions
and criticism.
